Quality of Service with Software Defined Network.

### 2017-03-07 report/paper.tex
delay extension

\textbf{Forwarding table:}f
\begin{itemize}
 \item \texttt{match fields:} sono i valori con cui i pacchetti vengono confrontati per cercare corrispondenze. Contengono la porta d'ingresso, l'intestazione del pacchetto, 
 e altri campi opzionali;
 \item \texttt{priority:} precedenza nel caso di voci con pi\`u corrispondenze;
 \item \texttt{counter:} incrementato ogni qualvolta che viene riscontrata la corrispondenza con un pacchetto;
 \item \texttt{istructions:} istruzioni da aggiungere all'\emph{Action-Set} o modifiche al flusso della pipeline;
 \item \texttt{timeouts:} massimo tempo per cui lo switch tiene in memoria questa voce, altre il quale la voce viene scartata;
 \item \texttt{cookie:} valori utilizzati per statistiche riguardo la singola voce, quindi non vengono considerati nell'elaborazione dei pacchetti dati;
 \item \texttt{flags:} utilizzate per differenziare la gestione della voce;
 \\
 \\
\end{itemize}

@@

\section{Instrument}\label{sec:instrument}
\#TODO matteo
  \subsection{Controller}
  The controller that we use is POX, the software is opensource and it is write with Python, it is based on SDN and support OpenFlow.
  The software use the version 2.7 of Python.
  It's repository is at this link \url{https://github.com/noxrepo/pox} and a lot of documentation about the progect
  is present at this link \url{https://openflow.stanford.edu/display/ONL/POX+Wiki}.
  The only problem to use this controller is that it support only the first version of OpenFlow1.0.
  \\
  The controller run inside a VirtualMachine on our personal computer,
  the VM is offered by SDN hub on this site: \url{http://sdnhub.org/tutorials/sdn-tutorial-vm/}.
  This VM is a 64-bit Ubuntu 14.04 that has a number of SDN software and tools installed.
  % \#TODO matteo da correggere e completare
  \subsection{Router Board}
  The router board that we used is MikroTik miniROUTER RouterBOARD 450G. This Router Board virtualize an OpenFlow switch that support
  OpenFlow v1.0\cite{routerboard_doc}.
  \begin{itemize}
   \item CPU		AR7161 680MHz
   \item Memory		256MB DDR SDRAM onboard memory
   \item Data storage	512MB onboard NAND memory chip, microSD card slot (on reverse)
   \item Ethernet 	Three 10/100/1000 Mbit/s Gigabit Ethernet ports supporting Auto-MDI/X
   \item Serial port 	One DB9 RS232C asynchronous serial port
  \end{itemize}
  \textbf{firmware} the current version is: '' da gercare col comando: system routerboard print ''
  \newline
  The \textbf{Operating System} that they use is MikroTik RouterOS ( cercare la versione installata) \cite{routeboard_software}.
  \newline
  One limitation of this routerboard is that it has only 5 ethernet port this is a limitation because consider that one of this port it is
  used to connect conntroller and with other port we have made host and other routerboard connection.

@@

\section{Results}\label{sec:results}
\#TODO %qui non so se serve un'introduzione o meno
  \subsection{Graph and Network structures}
  In order to achieve the realtime adaptability and statistics storage we have develop ad-hoc structure.
  We use a specific python library called NetworkX \cite{networkx} to create a graph in witch we can storage ad-hoc node that represent
  resplectively switch or host element.
  This topological structure is updated realtime thanks to connection event and link discover offered by various library.
  ( mettere i vari eventi che sfrutto )
  The link besides have many parameters in witch we can save network performance. To update every link parameter we have to creat ad hoc function
  because there isn't some function constructed yet. ( sistemare frase) .
  This type of structure allow us to run Dijkstra algorithm that return the minimum path ( in base a vari parametri ) on a nlog n time ( verificare).
  (aggiungere qualcosa e correggere)
  \#TODO matteo

  \subsection{Traffics analisis}\label{subsec:traf}
  For every networks performance we have tryed to develop a specific solution. The OpenFlow main goal is to centralize controll and software
  complexity so the routerboard can made only some easy function in orther to contain the hardware complexity and mantainly a good speed.
  Il che ha portato a dover utilizzare stratagemmi.
  \#TODO %riformulare
    \subsubsection{Delay}

	  Per molte applicazioni odierne il delay e' un parametro importantissimo.
    La rete attuale puo' conoscere il delay di un link semplicemente inviando un pachetto ICMP di tipo TimeStamp ( verificare )
    supportato dal protocollo openflow, questo permette di ottenere una precisione nell'ordine dei microsecondi ( verificare ).

    \#TODO MATTEO
		\subsubsection{Link Capacity}

		\#TODO andrea
		The first data that was collected was the actual link capacity, which depends both on the cable used to connect
		different devices and the effective capacity for each ethernet port linked by the cable itself. Everytime a device supporting
		openflow connects to the controller, it gives to the latter information about its status and its capabilities. One of them is
		the link capacity in three ways: the first is the \textit{supported} capacity, which is the phisical maximum capacity, the second is the
		\textit{advised} capacity, which name describes itself and the latter is the \textit{current} one. We choose to save the last one,
		as it gives the most useful about the link state. This information, moreover, is the one that non-SDN switches use to choose the \textit{metric}
		parameter. The simulation with Mininet showed us that this type of information can be used, but the Routeboards didn't support it.
		As a matter of fact, we weren't able to exploit this capacity in our network realization.
    \subsubsection{Throughput}

	  \#TODO andrea
		Using the flow statistics, we analyzed the traffic going through a link and, with a simple division, we got the normalized throughput.
		We exploited the information given by the switches with the match structure: this means that we used the throughput for the TCP traffic,
		the one for the UDP and for each of them we could know toward which IP the traffic is directed or is from. This was exploited with mininet
		simulations, but not in the set up in the laboratory. With the latter, we used the statistics as an indicator of the bitrate of a link
		and divided the host to categories based on it.

		\subsubsection{Packet error rate}

		The port statistics given by the switches contains different informations about the state of each port in a switch. To classify a link for its
		reliability, we divided the number of packets with erroneous symbols that couldn't be corrected with a FEC by the total number of packets that were sent,
		getting the packet error rate. The choice we made when exploiting this statistic was privileging the gaming and normal traffic, wherease the
		heavy traffic was directed to links with highest packet error rate.  As the switches exchanges a packet before recognizing most of the error, both because the
		phisical error are corrected using blocks coding and other error are found using a \textit{Frame Check Sequence} like CRC, the byte error were considered worse as
		those are indirect measures themself and the error derived from the calculations could grow up.
		\#TODO andrea

  \subsection{Route changing} %cambio rotte

@@

  \subsection{ARP solution} %TODO questione dell'arp inutile

  During our analysis and in our tests, we prove that in a SDN mesh network, the ARP packets are useless.
  This type of packets is useful for a multipoint to multipoint connection where the connection is set between a
  HUB and a set of computers or between a bridge and a computer. The first case is no more in use in common networks as
  hiher the collision range. The bridges limits their collision range to the port. With openflow each switching unit can
  act as a bridge or as a router. In both cases analyses the header of the L2 packet L3 packet and some of the L4 header.
  If it has the rule to forward the packet, it will act as declared in the most specific matching rule. Otherwise it will
  send to the controller a PacketIn and the controller will act as programmed. If an host want to communicate to another host whithin
  the network, it sends the packet to the line and then the switch will check in its tables and execute the actions of the match found.
  For an IP network, it will check the path to reach a destination and then send to the defined output port. The mac address in this case
  became useless, as the connection is point to point. So, knowing the destination layer2 address is useless as the switch can act
  regardless of it. The only thing the switch need to know is the layer2, layer 3 addresses and, at most, the layer 4 address.
  This can be seen by a reader as a violation of the layers in the ISO/OSI stack, but to exploit the potentials of the SDN approach,
  this is necessary. TODO explain why TODO!!!!**
  \#TODO andrea

@@

\begin{thebibliography}{99}
	\bibitem{pox} POX \url{https://openflow.stanford.edu/display/ONL/POX+Wiki}
	\bibitem{openflow} OpenFlow \url{http://flowgrammable.org/sdn/openflow/}
	\bibitem{pox_repo} POX's repository \url{https://github.com/noxrepo/pox}
	\bibitem{VM} Virtual Machine \url{http://sdnhub.org/tutorials/pox/}
	\bibitem{router_board} Router Board \url{https://www.senetic.it/product/RB450G}
	\bibitem{netoworkx} NetworkX python's library \url{https://networkx.github.io/}
	\bibitem{ONF} Open Networking Foundation \url{https://www.opennetworking.org/sdn-resources/technical-library}
	\bibitem{ONF_report} v1.5.1 OpenFlow report \url{https://www.opennetworking.org/sdn-resources/technical-library}
	\bibitem{routerboard_doc} RouterBOARD 450G MikroTik documentation \url{http://www.routerboard.sk/files/pdf/rb450g_manual.pdf}
	\bibitem{routeboard_software} RouterBOARD MikroTik Software \url{https://mikrotik.com/software}
	\bibitem{farzaneh}  Efficient Topology Discovery in Software Defined Networks F. Pakzad, M. Portmann, W. L. Tan and J. Indulska 
	School of ITEE, The University of Queensland Brisbane, Australia \url{https://www.researchgate.net/publication/271508903_Efficient_Topology_Discovery_in_OpenFlow-based_Software_Defined_Networks}
\end{thebibliography}

>>>>

\textbf{Forwarding table:}
\begin{itemize}
 \item \texttt{match fields:} sono i valori con cui i pacchetti vengono confrontati per cercare corrispondenze. Contengono la porta d'ingresso, l'intestazione del pacchetto, 
 e altri campi opzionali;
 \item \texttt{priority:} precedenza nel caso di voci con pi\`u corrispondenze;
 \item \texttt{counter:} incrementato ogni qualvolta che viene riscontrata la corrispondenza con un pacchetto;
 \item \texttt{istructions:} istruzioni da aggiungere all'\emph{Action-Set} o modifiche al flusso della pipeline;
 \item \texttt{timeouts:} massimo tempo per cui lo switch tiene in memoria questa voce, altre il quale la voce viene scartata;
 \item \texttt{cookie:} valori utilizzati per statistiche riguardo la singola voce, quindi non vengono considerati nell'elaborazione dei pacchetti dati;
 \item \texttt{flags:} utilizzate per differenziare la gestione della voce;
 \\
 \\
\end{itemize}

@@

\section{Instrument}\label{sec:instrument}
\#TODO matteo
  \subsection{Controller}
  The controller that we use is POX, the software is opensource and it is write with Python, it is based on SDN and support OpenFlow.
  The software use the version 2.7 of Python.
  It's repository is at this link \url{https://github.com/noxrepo/pox} and a lot of documentation about the progect
  is present at this link \url{https://openflow.stanford.edu/display/ONL/POX+Wiki}.
  The only problem to use this controller is that it support only the first version of OpenFlow1.0.
  \\
  The controller run inside a VirtualMachine on our personal computer,
  the VM is offered by SDN hub on this site: \url{http://sdnhub.org/tutorials/sdn-tutorial-vm/}.
  This VM is a 64-bit Ubuntu 14.04 that has a number of SDN software and tools installed.
  % \#TODO matteo da correggere e completare
  \subsection{Router Board}
  \begin{figure}[!h]
 \centering
 \includegraphics[scale=0.30]{images/rboard.png}
 \caption{\emph{Experience scenario}}
 \label{fig:topo}
  \end{figure}

  The router board that we used is MikroTik miniROUTER RouterBOARD 450G. This Router Board virtualize an OpenFlow switch that support
  OpenFlow v1.0\cite{routerboard_doc}.
  \begin{itemize}
   \item CPU		AR7161 680MHz
   \item Memory		256MB DDR SDRAM onboard memory
   \item Data storage	512MB onboard NAND memory chip, microSD card slot (on reverse)
   \item Ethernet 	Three 10/100/1000 Mbit/s Gigabit Ethernet ports supporting Auto-MDI/X
   \item Serial port 	One DB9 RS232C asynchronous serial port
  \end{itemize}
  \textbf{firmware} the current version is: '' da gercare col comando: system routerboard print ''
  \newline
  The \textbf{Operating System} that they use is MikroTik RouterOS ( cercare la versione installata) \cite{routeboard_software}.
  \newline
  One limitation of this routerboard is that it has only 5 ethernet port this is a limitation because consider that one of this port it is
  used to connect conntroller and with other port we have made host and other routerboard connection.

@@

\section{Results}\label{sec:results}
\subsection{Real Lab Scenario}
The topological scenario in which we have set up experience is the next:
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/topo.png}
 \caption{\emph{Experience scenario}}
 \label{fig:topo}
\end{figure}

We have used some principal function released with POX controller:
\begin{itemize}
 %\item \texttt{pox.topology.launch()} and \texttt{pox.openflow.topology.launch()} descrivere a che ci serve\cite{pox}.
 \item \texttt{pox.openflow.discovery.launch()} this function run algorithm in order to discovery the network and every time in witch some new 
 topological change occour it return an event with ( con cui ) we can update our topological knowledge.
 It does not provide the host event. For host event\cite{pox}.
 \item \texttt{pox.openflow.spanning_tree.launch()} this function is necessary with a presence of loop on topology 
 because with a spanning_tree algorithm it avoid the broadcast storm of message. When we set up \texttt{OFPP_FLOOD} like output port action in a table's 
 switch row if some spanning_tree algorithm run the \texttt{OFPP_FLOOD} is only theoretical infact the algorithm close some port for flooding service 
 on the switch in orther to avoid loop\cite{pox}.
\end{itemize}

\#TODO %qui non so se serve un'introduzione o meno
  \subsection{Graph and Network structures}
  In order to achieve the realtime adaptability and statistics storage we have develop ad-hoc structure.
  We use a specific python library called NetworkX \cite{networkx} to create a graph in witch we can storage ad-hoc node that represent
  resplectively switch or host element.
  This topological structure is updated realtime thanks to connection event and link discover offered by various library.
  ( mettere i vari eventi che sfrutto )
  The link besides have many parameters in witch we can save network performance. To update every link parameter we have to creat ad hoc function
  because there isn't some function constructed yet. ( sistemare frase) .
  This type of structure allow us to run Dijkstra algorithm that return the minimum path ( in base a vari parametri ) on a nlog n time ( verificare).
  (aggiungere qualcosa e correggere)
  \#TODO matteo

  \subsection{Traffics analisis}\label{subsec:traf}
  For every networks performance we have tryed to develop a specific solution. The OpenFlow main goal is to centralize controll and software
  complexity so the routerboard can made only some easy function in orther to contain the hardware complexity and mantainly a good speed\cite{qos_paper}.
  Il che ha portato a dover utilizzare stratagemmi.
  \#TODO %riformulare
    \subsubsection{Delay}
Monitoring round-trip time provides important insights for network troubleshooting and traffic engineering.
The  common  monitoring  technique  is  to  actively  send  probe packets  from  selected  vantage  points  (hosts  or  middleboxes).
In traditional  networks, the control over the network routing is
limited,  making  it  impossible  to  monitor  every  selected  path. The  emerging  concept  of  Software  Defined  Networking  sim-plifies 
network  control.  However,  OpenFlow,  the  common  SDN protocol,   does   not   support   RTT   monitoring   as   part   of   its
specification. In this experience we have tryed to emply an tecnic to escape this problem\cite{rtt}.
\newline The idea is to measure the time that occure from sending a packet from controller to a switch and the time in which the packet return from 
an other switch.
\\
\newline First we have to discovery round trip time from controller to every single switch on our topology.
In order to do this on the controller we start a timer, we send a proper packet to the switch specifing the \texttt{OFPP_CONTROLLER - Send to the controller}
option and when the message return to the controller we stop the timer and calculare RTT for the switch.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt2.png}
 \caption{\emph{RTT between controller and OF switch}}
 \label{fig:topo}
\end{figure}
\newline
If we need to calculate the delay on a specific link, the rtt has to calculate with the headed switch of this link.
\\
\newline 
Now we create a proper message with unique identifier number, we start a timer \texttt{TM1} and then send this message to a first switch specifing the output port.
Then when the packet comes to the second switch it identify the particular message and sends to the controller that stops it's dedicated timer \texttt{TM2}.
\newline This part required a particolar flow entry on the second switch in orther to send to the controller only some particulare packet.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt1.png}
 \caption{\emph{Pck from controller to itself following a particular way.}}
 \label{fig:topo}
\end{figure}
\\
\newline At this point we can calculate the delay of the link.
\\
\\
 delay = (TM2 - TM1) - RTT1/2 - RTT2/2
\\
\\
Teorethically this method must works but in practice there are a lot of delay introduce from computational concestion on the virtualized controller.
What we observe is a estimated delay like some seconds that is impossible, infact if we made some \texttt{ping test} from one host to another host 
in this SDN network we realize delay around 3ms.
\\

		\subsubsection{Link Capacity}

		\#TODO andrea
		The first data that was collected was the actual link capacity, which depends both on the cable used to connect
		different devices and the effective capacity for each ethernet port linked by the cable itself. Everytime a device supporting
		openflow connects to the controller, it gives to the latter information about its status and its capabilities. One of them is
		the link capacity in three ways: the first is the \textit{supported} capacity, which is the phisical maximum capacity, the second is the
		\textit{advised} capacity, which name describes itself and the latter is the \textit{current} one. We choose to save the last one,
		as it gives the most useful about the link state. This information, moreover, is the one that non-SDN switches use to choose the \textit{metric}
		parameter. The simulation with Mininet showed us that this type of information can be used, but the Routeboards didn't support it.
		As a matter of fact, we weren't able to exploit this capacity in our network realization.
		\\
		\\
    \subsubsection{Throughput}
	  \#TODO andrea
		Using the flow statistics, we analyzed the traffic going through a link and, with a simple division, we got the normalized throughput.
		We exploited the information given by the switches with the match structure: this means that we used the throughput for the TCP traffic,
		the one for the UDP and for each of them we could know toward which IP the traffic is directed or is from. This was exploited with mininet
		simulations, but not in the set up in the laboratory. With the latter, we used the statistics as an indicator of the bitrate of a link
		and divided the host to categories based on it.
\\
		\subsubsection{Packet error rate}

		The port statistics given by the switches contains different informations about the state of each port in a switch. To classify a link for its
		reliability, we divided the number of packets with erroneous symbols that couldn't be corrected with a FEC by the total number of packets that were sent,
		getting the packet error rate. The choice we made when exploiting this statistic was privileging the gaming and normal traffic, wherease the
		heavy traffic was directed to links with highest packet error rate.  As the switches exchanges a packet before recognizing most of the error, both because the
		phisical error are corrected using blocks coding and other error are found using a \textit{Frame Check Sequence} like CRC, the byte error were considered worse as
		those are indirect measures themself and the error derived from the calculations could grow up.
		\#TODO andrea
\\
  \subsection{Route changing} %cambio rotte

@@

  \subsection{ARP solution}

  During our analysis and in our tests, we prove that in a SDN mesh network, the ARP packets are useless.
  This type of packets is useful for a multipoint to multipoint connection where the connection is set between a
  HUB and a set of computers or between a bridge and a computer. The first case is no more in use in common networks as
  hiher the collision range. The bridges limits their collision range to the port. With openflow each switching unit can
  act as a bridge or as a router. In both cases analyses the header of the L2 packet L3 packet and some of the L4 header.
  If it has the rule to forward the packet, it will act as declared in the most specific matching rule. Otherwise it will
  send to the controller a PacketIn and the controller will act as programmed. If an host want to communicate to another host whithin
  the network, it sends the packet to the line and then the switch will check in its tables and execute the actions of the match found.
  For an IP network, it will check the path to reach a destination and then send to the defined output port. The mac address in this case
  became useless, as the connection is point to point. So, knowing the destination layer2 address is useless as the switch can act
  regardless of it. The only thing the switch need to know is the layer2, layer 3 addresses and, at most, the layer 4 address.
  This can be seen by a reader as a violation of the layers in the ISO/OSI stack, but to exploit the potentials of the SDN approach,
  this is necessary.

@@

\begin{thebibliography}{99}
	\bibitem{pox} POX \url{https://openflow.stanford.edu/display/ONL/POX+Wiki}
	\bibitem{openflow} OpenFlow \url{http://flowgrammable.org/sdn/openflow/}
	\bibitem{pox_repo} POX's repository \url{https://github.com/noxrepo/pox}
	\bibitem{VM} Virtual Machine \url{http://sdnhub.org/tutorials/pox/}
	\bibitem{router_board} Router Board \url{https://www.senetic.it/product/RB450G}
	\bibitem{networkx} NetworkX python's library \url{https://networkx.github.io/}
	\bibitem{ONF} Open Networking Foundation \url{https://www.opennetworking.org/sdn-resources/technical-library}
	\bibitem{ONF_report} v1.5.1 OpenFlow report \url{https://www.opennetworking.org/sdn-resources/technical-library}
	\bibitem{routerboard_doc} RouterBOARD 450G MikroTik documentation \url{http://www.routerboard.sk/files/pdf/rb450g_manual.pdf}
	\bibitem{routeboard_software} RouterBOARD MikroTik Software \url{https://mikrotik.com/software}
	\bibitem{farzaneh}  Efficient Topology Discovery in Software Defined Networks F. Pakzad, M. Portmann, W. L. Tan and J. Indulska 
	School of ITEE, The University of Queensland Brisbane, Australia \url{https://www.researchgate.net/publication/271508903_Efficient_Topology_Discovery_in_OpenFlow-based_Software_Defined_Networks}
	\bibitem{qos_paper} OpenNetMon: Network Monitoring in OpenFlow Software-Defined Networks Niels L.
	M. van Adrichem, Christian Doerr and Fernando A. Kuipers Network Architectures and Services, Delft University of Technology 
	Mekelweg 4, 2628 CD Delft, The Netherlands \url{https://pdfs.semanticscholar.org/ea94/7a9334484ee4c3bb6bcfb4d1590aece1b178.pdf}
	\bibitem{rtt} Efficient Round-Trip Time Monitoring in OpenFlow Networks A. Atary A. Bremler-Barr Interdisciplinary Center Herzliya \url{http://www.deepness-lab.org/pubs/infocom2016_grami.pdf}

	\end{thebibliography}
